The forestecology R package for fitting and assessing models of interspecies competitive effects on the growth of trees

### 2021-03-01 paper/forestecology.Rmd
Incorporated @simonpcouch comments

The output `growth_scbi` is a single data frame of class `sf` that includes variables `growth`, the average annual growth in DBH (cm y^-1^) for all stems that were alive at both time points, and `geometry`, the `sf` package's encoding of geolocations of type `<POINT>`. Furthermore, the variables that should remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables. Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`. The data format of other sites may be such that our `compute_growth()` function doesn't work at all. However, in the end all that matters is that the growth of all stems is saved in a data frame of class `sf` and at a minimum the data contains the variables above. TODO: enumerate what these variables are and say if they need exact names

@@

Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest: all trees within a distance `comp_dist` of a focal tree are considered its competitors (assuming the same units as the `gx` and `gy` location variables). In our case we set this value at 7.5m, a value informed by @canham_neighborhood_2004 @uriarte_spatially_2004 @canham_neighborhood_2006. Using this value along with a manually constructed `sf` object representation of the study region's boundary, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not.

@@

The next step in our analysis sequence is to identify all focal trees and their corresponding competitor trees. The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf`. On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds; in our case, this was manually encoded in the `blocks_scbi` in Section \ref{spatial-information} while in our Appendix we present an example where this was performed using `spatialBlock()` from the `blockCV` package. We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.

@@

The resulting data frame `focal_vs_comp_scbi` has `r nrow(focal_vs_comp_scbi)` rows, representing the subset of the `r nrow(growth_scbi)` trees in `growth_scbi` that will be considered as focal trees. Two new variables `focal_ID` and `focal_sp` relate to tree-stem identification and species information. Most notably however is a new variable `comp` which contains information on all competitor trees for a given focal tree, saved in `tidyr` package list-column format @tidyr_package. For example, we drill-down on the tree with `focal_ID` 4, which has 20 competitor trees each described by 4 variables as indicated by the fact that `comp` is a `<tibble [20 × 4]>`.

@@

Now that we've identified all focal and corresponding competitor trees and saved this information in a data frame of type `focal_vs_comp`, the final step in our analysis sequence is to fit a model for the growth of all focal trees. Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Section \ref{competition-model} using the `comp_bayes_lm()` function. However, any model implemented in a function that similarly takes an input data frame of type `focal_vs_comp` as an argument can be easily swapped in. For our specific competition Bayesian linear regression model, we also specify prior distributions on all parameters of interest (here chosen to be the defaults as specified in `?comp_bayes_lm`).

>>>>

The output `growth_scbi` is a single data frame of class `sf` that includes variables `growth`, the average annual growth in DBH (cm y^-1^) for all stems that were alive at both time points, and `geometry`, the `sf` package's encoding of geolocations of type `<POINT>`. Furthermore, the variables that should remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables. Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`. Site data that does not align with this convention will need to be transformed for use with the `compute_growth()` function. However, in the end all that matters is that the growth of all stems is saved in a data frame of class `sf` and at a minimum the data contains the variables above. TODO: enumerate what these variables are and say if they need exact names

@@

Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest: all trees within a distance `comp_dist` of a focal tree are considered its competitors (assuming the same units as the `gx` and `gy` location variables). In our case we set this value at 7.5m, a value informed by @canham_neighborhood_2004 @uriarte_spatially_2004 @canham_neighborhood_2006. Using this value along with a manually constructed `sf` object representation of the study region's boundary via its vertices, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not.

@@

The next step in our analysis sequence is to identify all focal trees and their corresponding competitor trees. The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf`. On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds. In this example, the blocks were manually encoded in `blocks_scbi` by specifying it's vertices in Section \ref{spatial-information}^[We present an alternative method for defining spatial cross-validation blocks is using the `spatialBlock()` function from the `blockCV` package in the Appendix]. We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.

@@

The resulting data frame `focal_vs_comp_scbi` has `r nrow(focal_vs_comp_scbi)` rows, representing the subset of the `r nrow(growth_scbi)` trees in `growth_scbi` that will be considered as focal trees. Two new variables `focal_ID` and `focal_sp` relate to tree-stem identification and species information. Most notably however is a new variable `comp` which contains information on all competitor trees for a given focal tree, saved in `tidyr` package list-column format @tidyr_package. For example, we drill-down on the tree with `focal_ID` 4, which has 20 competitor trees each described by 4 variables as indicated by the fact that `comp` is a `tibble [20 × 4]`.

@@

Now that we've identified all focal and corresponding competitor trees and saved this information in a data frame of type `focal_vs_comp`, the final step in our analysis sequence is to fit a model for the growth of all focal trees. Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Section \ref{competition-model} using the `comp_bayes_lm()` function. However, any model implemented in a function that similarly takes an input data frame of type `focal_vs_comp` as an argument can also be used. For our specific competition Bayesian linear regression model, we also specify prior distributions on all parameters of interest (here chosen to be the defaults as specified in `?comp_bayes_lm`).
